19Q1

### 2019-04-05 19Q1.Rmd
created 19Q1.Rmd for htscreens work

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached):å
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

```{r}
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(reshape2)
library(tidyverse)
library(VennDiagram)
theme_set(theme_light())
```

# DATA MANAGEMENT {.tabset .tabset-fade .tabset-pills}

@@

- For their universal ccancer cell line nomenclature, they reverted back to `Broad_ID` from 18Q3 rather than `DepMap_ID` that was used in 18Q4
- The `Primary.Disease` column of the cell line metadata file is more consistently named (`Subtype.Disease` now has the specific disease annotations), though there are a couple capitalization inconsistencies
- MAF file no longer has `CCLE_Name` column and has a duplicate column of the Broad IDs (i.e. `DepMap_ID` and `Tumor_Sample_Barcode`)

@@

## DepMap cell line metadata

19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_info <- read.delim("./data_munging/19Q1/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$source <- "CCLE"
ccl_info$source_ccl_name <- strsplit(as.character(ccl_info$CCLE_Name), split = "_")
ccl_info$source_ccl_name <- as.character(lapply(ccl_info$source_ccl_name, function(x) ifelse("[MERGED" %in% x || "NORMAL" %in% x, NA, x[1])))
ccl_info$source_ccl_name[ccl_info$source_ccl_name == "NA"] <- NA
ccl_info$source_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_info$CCLE_Name, ignore.case = TRUE))
ccl_info$source_histology <- ifelse(ccl_info$CCLE_Name == ccl_info$source_ccl_name, NA, ccl_info$source_histology)
ccl_info$sample_ID <- with(ccl_info, paste(source, source_ccl_name, source_histology, sep = "_"))
# This CCL had a histology of "primary": Looking at the DepMap reference, this is an immortalize epithelial line from the CNS (https://depmap.org/portal/cell_line/NHAHTDD_PRIMARY), so I chanegd the histology accordingly
ccl_info$source_histology[ccl_info$source_ccl_name == "NHAHTDD"] <- "central_nervous_system"
# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_info$Primary.Disease <- as.factor(tolower(ccl_info$Primary.Disease))
# Harmonize ID column name and rename CCLE_Name column to fit with dataframe style
colnames(ccl_info)[colnames(ccl_info) == "CCLE_Name"] <- "CCLE_ccl_name"
colnames(ccl_info)[colnames(ccl_info) == "DepMap_ID"] <- "Broad_ID"

# Sample info from figshare
crispr_meta <- read.delim("./data_munging/19Q1/sample_info_20190306.csv", sep = ",", header = TRUE, na.strings = c("", NA))
colnames(crispr_meta)[colnames(crispr_meta) == "CCLE_name"] <- "CCLE_Name"
```

@@

@@

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_corrected_20190305.csv.gz", sep = ",", check.names = FALSE, header = TRUE)
```

# Pozdeyev et al. (2016) summary

```{r}
ccl_histologies <- select(ccl_info, source_ccl_name, source_histology, CCLE_ccl_name, Broad_ID) %>% drop_na(source_ccl_name)
colnames(ccl_histologies)[colnames(ccl_histologies) == "source_ccl_name"] <- "Cell_line" 
colnames(ccl_histologies)[colnames(ccl_histologies) == "source_histology"] <- "Histology"

# Account for cell line aliases and missing histologies
auc_na_histologies <- read.delim("./data_munging/pozdeyev2016/auc_na_histologies.csv", sep = ",", row.names = 1)

auc_na_histologies$accession_id <- NULL
auc_na_histologies$Notes <- NULL
auc_na_histologies$Histology <- ifelse(auc_na_histologies$Histology == "", NA, as.character(auc_na_histologies$Histology))
auc_na_histologies$CCLE_ccl_name <- ifelse(auc_na_histologies$CCLE_ccl_name  == "", NA, as.character(auc_na_histologies$CCLE_ccl_name ))
auc_na_histologies$Broad_ID <- ifelse(auc_na_histologies$Broad_ID == "", NA, as.character(auc_na_histologies$Broad_ID))

ccl_histologies <- subset(ccl_histologies, !(Broad_ID %in% auc_na_histologies$Broad_ID))
# Bind DepMap histologies and curated histologies
ccl_histologies <- bind_rows(ccl_histologies, auc_na_histologies)
```

For plotting:

- `c("CCLE" = "springgreen4", "CTRP" = "steelblue4", "GDSC" = "turquoise4")`
- `c("springgreen4", "steelblue4", "turquoise4")`

## Data wrangling

```{r, eval=FALSE}
ccle_poz <- read.delim("./data_munging/pozdeyev2016/CCLE_drug_response_metrics.csv.gz", sep = ",")

# ccle_test <- merge(select(ccle_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CCLE"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

ctrp_poz <- read.delim("./../htscreens_giant_files/pozdeyev2016/CTRP_drug_response_metrics.csv.gz", sep = ",")
# ctrp_test <- merge(select(ctrp_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "CTRP"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

gdsc_poz <- read.delim("./data_munging/pozdeyev2016/GDSC_drug_response_metrics.csv.gz", sep = ",")
# gdsc_test <- merge(select(gdsc_poz, Drug_name, Cell_line, AUC_IC50), select(filter(auc, Database == "GDSC"), Drug_name, Cell_line, AUC_IC50), by = c("Drug_name", "Cell_line"), suffixes = c("_unmatched", "_matched_all_pairs"), all = TRUE)

auc <- bind_rows(ccle_poz, ctrp_poz, gdsc_poz)

>>>>

```{r}
library(crunch)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(kableExtra)
library(naniar)
library(openxlsx)
library(reshape2)
library(tidyverse)
library(VennDiagram)

theme_set(theme_light())
```

# DATA

@@

- For their universal ccancer cell line nomenclature, they reverted back to `Broad_ID` from 18Q3 rather than `DepMap_ID` that was used in 19Q1
- The `Primary.Disease` column of the cell line metadata file is more consistently named (`Subtype.Disease` now has the specific disease annotations), though there are a couple capitalization inconsistencies
- MAF file no longer has `CCLE_Name` column and has a duplicate column of the Broad IDs (i.e. `DepMap_ID` and `Tumor_Sample_Barcode`)

@@

## Cell line metadata

DepMap 19Q1 update fixed almost all of the inconsistencies of the Primary.Disease column, though there were come capitalization inconsistencies.
```{r}
ccl_info <- read.delim("./data_munging/19Q1/DepMap-2019q1-celllines_v2_20190305.csv", sep = ",", header = TRUE, na.strings = c("", NA))
ccl_info$source <- "CCLE"
ccl_info$Cell_line <- strsplit(as.character(ccl_info$CCLE_Name), split = "_")
ccl_info$Cell_line <- as.character(lapply(ccl_info$Cell_line, function(x) ifelse("[MERGED" %in% x || "NORMAL" %in% x, NA, x[1])))
ccl_info$Cell_line[ccl_info$Cell_line == "NA"] <- NA
ccl_info$source_histology <- tolower(gsub("^[a-z0-9]*_", "", ccl_info$CCLE_Name, ignore.case = TRUE))
ccl_info$source_histology <- ifelse(ccl_info$CCLE_Name == ccl_info$CCLE_Name, NA, ccl_info$source_histology)

# This CCL had a histology of "primary": Looking at the DepMap reference, this is an immortalize epithelial line from the CNS (https://depmap.org/portal/cell_line/NHAHTDD_PRIMARY), so I chanegd the histology accordingly
ccl_info$source_histology[ccl_info$Cell_line == "NHAHTDD"] <- "central_nervous_system"
# Convert Primary.Disease entries to all lower case to fix factorization issue
ccl_info$Primary.Disease <- as.factor(tolower(ccl_info$Primary.Disease))
# Harmonize ID column name and rename CCLE_Name column to fit with dataframe style
colnames(ccl_info)[colnames(ccl_info) == "CCLE_Name"] <- "CCLE_ccl_name"
colnames(ccl_info)[colnames(ccl_info) == "DepMap_ID"] <- "Broad_ID"
```

Adding TCGA labels:
```{r, eval=FALSE}
tcga_hist_converter <- read.xlsx("./data_munging/TCGA/cell_line_annotation.xlsx", sheet = 1, startRow = 1, colNames = TRUE, rowNames = FALSE)
tcga_hist_converter <- add_column(.data = tcga_hist_converter, TCGA_manual = NA, .after = "TCGA_combine")
tcga_hist_converter <- tcga_hist_converter[with(tcga_hist_converter, order(primary_DepMap, subtype_DepMap, primary_site_CTRP, primary_hist_CTRP, subtype_CTRP, tissue1_GDSC)),]

# For manual curation of TCGA labels
# ccls_for_manual_tcga_curation <- filter(tcga_hist_converter, is.na(TCGA_combine))
# ccls_for_manual_tcga_curation <- ccls_for_manual_tcga_curation[with(ccls_for_manual_tcga_curation, order(primary_DepMap, subtype_DepMap, primary_site_CTRP, primary_hist_CTRP, subtype_CTRP, tissue1_GDSC)),]
# write.csv(ccls_for_manual_tcga_curation, "./data_munging/TCGA/cell_line_annotation_noTCGA.csv")
tcga_hist_converter <- filter(tcga_hist_converter, !is.na(TCGA_combine))
ccls_curated <- read.delim("./data_munging/TCGA/cell_line_annotation_noTCGA_curated.csv", sep = ",", row.names = 1)
tcga_histsconverter <- bind_rows(tcga_hist_converter, ccls_curated)
tcga_hist_converter$TCGA_combine <- with(tcga_hist_converter, ifelse(is.na(TCGA_combine), as.character(TCGA_manual), TCGA_combine))
# write.csv(tcga_hist_converter, "./data_munging/TCGA/cell_line_annotation_curated_sally_20190404.csv")
```

## G2P

```{r}
g2p <- read.delim("./data_munging/g2p_full_level_A_dataset_NEW_DEC_2_2018_withGroupedLineages.csv", sep = "\t", header = TRUE)
g2p[, 1:2] <- NULL
g2p <- merge(g2p, lineage_converter, by = "grouped_general_doid", all.x = TRUE)

g2p$Drug <- tolower(g2p$Drug)
g2p[g2p == -1] <- NA
g2p$Ref[g2p$Ref == "None"] <- NA
g2p$Alt[g2p$Alt == "None"] <- NA
g2p$Drug_Gene <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene), NA, paste0(g2p$Drug, "_", g2p$Gene))
g2p$Drug_Gene_Lin <- ifelse(is.na(g2p$Drug) | is.na(g2p$Gene) | is.na(g2p$group_general_lineage_name), NA, paste0(g2p$Drug, "_", g2p$Gene, "_", g2p$group_general_lineage_name))
g2p$position <- paste0("g.chr", g2p$Chromosome, ":", g2p$End, g2p$Ref, ">", g2p$Alt)
g2p$position <- ifelse(grepl("NA", g2p$position), NA, g2p$position)
g2p_filt <- subset(g2p, grepl("^CID", g2p$DrugID))

g2p_dgls <-  unique(g2p_filt$Drug_Gene_Lin[!is.na(g2p_filt$Drug_Gene_Lin)])
g2p_druggenes <- unique(g2p_filt$Drug_Gene[!is.na(g2p_filt$Drug_Gene)])
g2p_drugs <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[1]))
g2p_genes <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[2]))
g2p_lins <- unique(sapply(strsplit(g2p_dgls, "_", fixed = TRUE), function(x) x[3]))
g2p_druggenes
g2p_drugs
g2p_genes
g2p_lins

# Order and genes taken from pink heatmapo from Daniel's analyses
g2p_genes <- c("EGFR", "KRAS", "ERBB2", "ERBB3", "MET", "ALK", "ROS1", "PDGFRA", "ABL1", "KIT", "PDGFRB", "PDGFB", "BRAF", "G6PD", "RET", "BRCA1", "BRCA2", "UGT1A1", "TSC1", "TSC2", "AKT1", "IDH2", "DPYD", "ESR1", "CYP19A1", "TPMT", "FLT3", "NPM1", "DNMT3A", "PML", "PTCH1", "JAK2", "MYD88", "MGMT")
```

```{r}
# levels(g2p_filt$Phenotype.Description)
unique(g2p_filt$group_general_lineage_name)
levels(g2p_filt$group_general_lineage_name)

g2p_nonCID <- subset(g2p, !grepl("^CID", g2p$DrugID))
summ <- g2p_nonCID %>% group_by(Drug, DrugID) %>% tally() %>% summarize(DrugID = DrugID, n = n, percent = formatC(n / 402 * 100, digits = 6, format = "f"))
summ
```

@@

## Drug screens

### Merge drug data and annotations

```{r}
test <- filter(maf_df, Hugo_Symbol %in% g2p_genes) %>% count(Hugo_Symbol, CCLE_Name)
test <- test[order(test$n, decreasing = TRUE),]
```

`r nrow(filter(test, n == 1))`/`r nrow(test)` gene-cell line combinations have more than one mutation in the MAF file. This is corrected in the point mutation filtering done below.

```{r, eval=FALSE}
# Drug metadata
drug_meta <- read.delim("./data_munging/drug_id_drug_name_table_10_12_2018.tsv", sep = "\t", header = FALSE)
colnames(drug_meta) <- c("CID", "Drug")
drug_meta$Drug <- tolower(drug_meta$Drug)
maf_df_g2p <- filter(maf_df, Hugo_Symbol %in% g2p_genes)
ge_filt_18Q3 <- filter(ge_melt_18Q3, Hugo_Symbol %in% g2p_genes)
# ge_filt_19Q1 <- filter(ge_melt_19Q1, Hugo_Symbol %in% g2p_genes)
cn_filt <- filter(cn_melt, Hugo_Symbol %in% g2p_genes)

# CCLE
ccle <- read.delim("./data_munging/data_drug_auc_ccle_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ccle <- merge(drug_meta, ccle, by = "CID")
ccle_melt <- melt(ccle, id.vars = c("CID", "Drug"), measure.vars = colnames(ccle)[3:ncol(ccle)], variable.name = "accession_id", value.name = "ActArea")
ccle_grid <- expand.grid("accession_id" = unique(ccle_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ccle_data <- merge(ccle_grid, ccle_melt, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, ccl_converter, by = "accession_id", all.x = TRUE)
ccle_data <- merge(ccle_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ccle_data$Mutation_Status_Nonsilent <- ifelse(is.na(ccle_data$Mutation_Status_Nonsilent), "Wildtype", ccle_data$Mutation_Status_Nonsilent)
ccle_data <- merge(ccle_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ccle_data <- merge(ccle_data, ge_filt_19Q1, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data <- merge(ccle_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ccle_data$Hugo_Symbol <- factor(ccle_data$Hugo_Symbol)
ccle_data$ActArea <- as.numeric(ccle_data$ActArea)
# CCLE AUC is actually AOC
# assuming scale is 0-8, transform to range between 0 and 1
# get AUC by 1 - AOC
ccle_data$AUC <- 1 - (ccle_data$ActArea / 8)
ccle_data <- filter(ccle_data, !is.na(ccle_data$AUC))
ccle_data$AUC_zscore <- (ccle_data$AUC - mean(ccle_data$AUC)) / sd(ccle_data$AUC)
ccle_data$Color_Nonsilent <- ifelse(ccle_data$Mutation_Status_Nonsilent == "Wildtype", "palegreen3", "springgreen4")
ccle_data$Color_Nonsilent <- factor(ccle_data$Color_Nonsilent)
ccle_data$Drug_Gene <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol)
ccle_data$Drug_Gene_Lin <- paste0(ccle_data$Drug, "_", ccle_data$Hugo_Symbol, "_", ccle_data$group_general_lineage_name)
saveRDS(ccle_data, "./data_munging/rds/ccle_data_19Q1_g2pgenes.rds", compress = "xz")

# CTRP
ctrp <- read.delim("./data_munging/data_drug_auc_ctrp_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
ctrp <- merge(drug_meta, ctrp, by = "CID")
ctrp_melt <- melt(ctrp, id.vars = c("CID", "Drug"), measure.vars = colnames(ctrp)[3:ncol(ctrp)], variable.name = "accession_id", value.name = "AUC")
ctrp_grid <- expand.grid("accession_id" = unique(ctrp_melt$accession_id), "Hugo_Symbol" = g2p_genes)
ctrp_data <- merge(ctrp_grid, ctrp_melt, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, ccl_converter, by = "accession_id", all.x = TRUE)
ctrp_data <- merge(ctrp_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
ctrp_data$Mutation_Status_Nonsilent <- ifelse(is.na(ctrp_data$Mutation_Status_Nonsilent), "Wildtype", ctrp_data$Mutation_Status_Nonsilent)
ctrp_data <- merge(ctrp_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# ctrp_data <- merge(ctrp_data, ge_filt_19Q1, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data <- merge(ctrp_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
ctrp_data$Hugo_Symbol <- factor(ctrp_data$Hugo_Symbol)
ctrp_data$AUC <- as.numeric(ctrp_data$AUC)
ctrp_data <- filter(ctrp_data, !is.na(ctrp_data$AUC))
ctrp_data$AUC_zscore <- (ctrp_data$AUC - mean(ctrp_data$AUC)) / sd(ctrp_data$AUC)
ctrp_data$Color_Nonsilent <- ifelse(ctrp_data$Mutation_Status_Nonsilent == "Wildtype", "slategray3", "steelblue4")
ctrp_data$Color_Nonsilent <- factor(ctrp_data$Color_Nonsilent)
ctrp_data$Drug_Gene <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol)
ctrp_data$Drug_Gene_Lin <- paste0(ctrp_data$Drug, "_", ctrp_data$Hugo_Symbol, "_", ctrp_data$group_general_lineage_name)
saveRDS(ctrp_data, "./data_munging/rds/ctrp_data_19Q1_g2pgenes.rds", compress = "xz")

# GDSC
gdsc <- read.delim("./data_munging/data_drug_auc_gdsc_10_12_2018.csv", sep = "\t", header  = TRUE, row.names = 1, check.names = FALSE)
gdsc <- merge(drug_meta, gdsc, by = "CID")
gdsc_melt <- melt(gdsc, id.vars = c("CID", "Drug"), measure.vars = colnames(gdsc)[3:ncol(gdsc)], variable.name = "accession_id", value.name = "AUC")
gdsc_grid <- expand.grid("accession_id" = unique(gdsc_melt$accession_id), "Hugo_Symbol" = g2p_genes)
gdsc_data <- merge(gdsc_grid, gdsc_melt, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, ccl_converter, by = "accession_id", all.x = TRUE)
gdsc_data <- merge(gdsc_data, maf_df_g2p, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
gdsc_data$Mutation_Status_Nonsilent <- ifelse(is.na(gdsc_data$Mutation_Status_Nonsilent), "Wildtype", gdsc_data$Mutation_Status_Nonsilent)
gdsc_data <- merge(gdsc_data, ge_filt_18Q3, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
# gdsc_data <- merge(gdsc_data, ge_filt_19Q1, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
gdsc_data <- merge(gdsc_data, cn_filt, by = c("Hugo_Symbol", "DepMap_ID"), all.x = TRUE)
gdsc_data$Hugo_Symbol <- factor(gdsc_data$Hugo_Symbol)
gdsc_data$AUC <- as.numeric(gdsc_data$AUC)
gdsc_data <- filter(gdsc_data, !is.na(gdsc_data$AUC))
gdsc_data$AUC_zscore <- (gdsc_data$AUC - mean(gdsc_data$AUC)) / sd(gdsc_data$AUC)
gdsc_data$Color_Nonsilent <- ifelse(gdsc_data$Mutation_Status_Nonsilent == "Wildtype", "paleturquoise3", "turquoise4")
gdsc_data$Color_Nonsilent <- factor(gdsc_data$Color_Nonsilent)
gdsc_data$Drug_Gene <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol)
gdsc_data$Drug_Gene_Lin <- paste0(gdsc_data$Drug, "_", gdsc_data$Hugo_Symbol, "_", gdsc_data$group_general_lineage_name)
saveRDS(gdsc_data, "./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds", compress = "xz")
```

### Filter drug data

```{r, eval=FALSE}
ccle_data <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ccle_data_ptmuts <- filter(ccle_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
ccle_data_ptmuts$Keep_PtMut_Tests <- ifelse(ccle_data_ptmuts$Reference_Allele == 1 | is.na(ccle_data_ptmuts$Reference_Allele), TRUE, ifelse(ccle_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
ccle_data_ptmuts <- filter(ccle_data_ptmuts, Keep_PtMut_Tests == TRUE)
# Remove columns that lead to duplicate mutation entries
ccle_data_ptmuts <- subset(ccle_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
ccle_data_ptmuts$Dataset <- "CCLE"
ccle_data_ptmuts$Hugo_Symbol <- ordered(ccle_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
ccle_data_ptmuts <- unique(ccle_data_ptmuts)
saveRDS(ccle_data_ptmuts, "./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds", compress = "xz")

ctrp_data <- readRDS("./data_munging/rds/ctrp_data_19Q1_g2pgenes.rds")
ctrp_data_ptmuts <- filter(ctrp_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
ctrp_data_ptmuts$Keep_PtMut_Tests <- ifelse(ctrp_data_ptmuts$Reference_Allele == 1 | is.na(ctrp_data_ptmuts$Reference_Allele), TRUE, ifelse(ctrp_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
ctrp_data_ptmuts <- filter(ctrp_data_ptmuts, Keep_PtMut_Tests == TRUE)
# Remove columns that lead to duplicate mutation entries
ctrp_data_ptmuts <- subset(ctrp_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
ctrp_data_ptmuts$Dataset <- "CTRP"
ctrp_data_ptmuts$Hugo_Symbol <- ordered(ctrp_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
ctrp_data_ptmuts <- unique(ctrp_data_ptmuts)
saveRDS(ctrp_data_ptmuts, "./data_munging/rds/ctrp_data_19Q1_g2pgenes_ptmuts.rds", compress = "xz")

gdsc_data <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
gdsc_data_ptmuts <- filter(gdsc_data, Reference_Allele == 1 | Reference_Allele == 0 | is.na(Reference_Allele))
gdsc_data_ptmuts$Keep_PtMut_Tests <- ifelse(gdsc_data_ptmuts$Reference_Allele == 1 | is.na(gdsc_data_ptmuts$Reference_Allele), TRUE, ifelse(gdsc_data_ptmuts$Tumor_Seq_Allele1 == 1, TRUE, FALSE))
gdsc_data_ptmuts <- filter(gdsc_data_ptmuts, Keep_PtMut_Tests == TRUE)
# Remove columns that lead to duplicate mutation entries
gdsc_data_ptmuts <- subset(gdsc_data_ptmuts, select = -c(Variant_Classification, Reference_Allele, Tumor_Seq_Allele1, Genome_Change, Keep_PtMut_Tests))
gdsc_data_ptmuts$Dataset <- "GDSC"
gdsc_data_ptmuts$Hugo_Symbol <- ordered(gdsc_data_ptmuts$Hugo_Symbol, levels = g2p_genes)
gdsc_data_ptmuts <- unique(gdsc_data_ptmuts)
saveRDS(gdsc_data_ptmuts, "./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds", compress = "xz")
```

```{r}
ccle_data <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes.rds")
ctrp_data <- readRDS("./data_munging/rds/ctrp_data_19Q1_g2pgenes.rds")
gdsc_data <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes.rds")
ccle_data_ptmuts <- readRDS("./data_munging/rds/ccle_data_19Q1_g2pgenes_ptmuts.rds")
ctrp_data_ptmuts <- readRDS("./data_munging/rds/ctrp_data_19Q1_g2pgenes_ptmuts.rds")
gdsc_data_ptmuts <- readRDS("./data_munging/rds/gdsc_data_19Q1_g2pgenes_ptmuts.rds")

# Filter drug screen datasets for G2P drug-gene associations
ccle_data_g2p <- filter(ccle_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ccle_data_g2p$Mutation_Status_Nonsilent <- factor(ccle_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p <- filter(ctrp_data_ptmuts, Drug_Gene %in% g2p_druggenes)
ctrp_data_g2p$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p <- filter(gdsc_data_ptmuts, Drug_Gene %in% g2p_druggenes)
gdsc_data_g2p$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drug-gene-lineage associations
ccle_data_g2p_dgl <- filter(ccle_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ccle_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_dgl <- filter(ctrp_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
ctrp_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_dgl <- filter(gdsc_data_ptmuts, Drug_Gene_Lin %in% g2p_dgls)
gdsc_data_g2p_dgl$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_dgl$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Filter drug screen datasets for G2P drugs and genes
ccle_data_g2p_grid <- filter(ccle_data_ptmuts, Drug %in% g2p_drugs)
ccle_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ccle_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
ctrp_data_g2p_grid <- filter(ctrp_data_ptmuts, Drug %in% g2p_drugs)
ctrp_data_g2p_grid$Mutation_Status_Nonsilent <- factor(ctrp_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))
gdsc_data_g2p_grid <- filter(gdsc_data_ptmuts, Drug %in% g2p_drugs)
gdsc_data_g2p_grid$Mutation_Status_Nonsilent <- factor(gdsc_data_g2p_grid$Mutation_Status_Nonsilent, levels = c("Mutant", "Wildtype"))

# Testing
test_ccle <- ccle_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_ctrp <- ctrp_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()
test_gdsc <- gdsc_data_ptmuts %>% group_by(Hugo_Symbol, Mutation_Status_Nonsilent) %>% tally()

# drug_summ <- data.frame(Drug = g2p_drugs, Dataset_Count = 0)
```

How many datasets screen each G2P drug?
```{r}
drug_summ <- data.frame(Drug = character(), Dataset_Count = integer(), stringsAsFactors = FALSE)

for(drug in g2p_drugs) {
  counter = 0
  if(drug %in% as.character(unique(ccle_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(ctrp_data_g2p$Drug))) { counter <- counter + 1 }
  if(drug %in% as.character(unique(gdsc_data_g2p$Drug))) { counter <- counter + 1 }
  drug_summ <- rbind(drug_summ, data.frame(Drug = drug, Dataset_Count = counter))
}
```

@@

### DepMap dependency probabilities

```{r, eval=FALSE}
dep <- read.delim("./data_munging/19Q1/gene_dependency_19Q1_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(dep) <- gsub(" .*", "", colnames(dep))
```

### Merge CRISPR data and annotations

```{r, eval=FALSE}
crispr <- read.delim("./data_munging/19Q1/gene_effect_19Q1_20181115.csv.gz", sep = ",", header  = TRUE, check.names = FALSE)
# Remove Entrez gene IDs from colnames
colnames(crispr) <- gsub(" .*", "", colnames(crispr))
```

Merge annotation data:
```{r, eval=FALSE}
# Melt CRISPR dataset for merging
crispr_melt <- melt(crispr, id.vars = "DepMap_ID", measure.vars = colnames(crispr)[2:ncol(crispr)], variable.name = "Hugo_Symbol", value.name = "Score")
# Melt dependency probabilities dataset for merging
dep_melt <- melt(dep, id.vars = "DepMap_ID", measure.vars = colnames(dep)[2:ncol(dep)], variable.name = "Hugo_Symbol", value.name = "Dep_Prob")
# Merge dependency probabilities
crispr_melt <- merge(crispr_melt, dep_melt, by = c("DepMap_ID", "Hugo_Symbol"), all.x = TRUE)
# Merge cell line metadata
crispr_melt <- merge(crispr_melt, ccl_info, by = "DepMap_ID", all.x = TRUE)
crispr_melt <- merge(crispr_melt, crispr_meta, by = c("CCLE_Name", "DepMap_ID"), all.x = TRUE)

# Merge mutation annotations
maf_filt_crispr <- filter(maf_df, Hugo_Symbol %in% as.character(unique(crispr_melt$Hugo_Symbol)))
crispr_muts <- merge(crispr_melt, maf_filt_crispr, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
crispr_muts$Hugo_Symbol <- factor(crispr_muts$Hugo_Symbol)
crispr_muts <- crispr_muts %>% mutate(Mutation_Status_Nonsilent = if_else(is.na(Mutation_Status_Nonsilent), "Wildtype", Mutation_Status_Nonsilent))
# Summarize number of mutant and Wildtype cell lines
crispr_muts_summ <- crispr_muts %>% group_by(Hugo_Symbol) %>%
  summarize(N_Nonsilent_Wildtype = sum(Mutation_Status_Nonsilent == "Wildtype"),
            N_Nonsilent_Mutant = sum(Mutation_Status_Nonsilent == "Mutant"))
# Merge test results back into full dataset, which restores information lost in the summarization
crispr_data <- merge(crispr_muts_summ, crispr_muts, by = "Hugo_Symbol")
# Add Color columns
crispr_data$Color_Nonsilent <- ifelse(crispr_data$Mutation_Status_Nonsilent == "Wildtype", "thistle", "darkorchid")
crispr_data$Color_Nonsilent <- factor(crispr_data$Color_Nonsilent)
